Tree height, microhabitat, and hydraulic traits shape drought responses in a temperate broadleaf forest

### 2019-09-27 manuscript/climate_sensitivity_draft.Rmd
updated with Stovall reference

\newpage
### Introduction
Forests globally play a critical role in climate regulation ([Bonan 2008](https://science.sciencemag.org/content/320/5882/1444)), yet there remains enormous uncertainty as to how the terrestrial carbon (C) sink, which is dominated by forests, will respond to climate change ([Friedlingstein et al. 2006](https://journals.ametsoc.org/doi/full/10.1175/JCLI3800.1)). An important aspect of this uncertainty lies in responses to drought (**REF**). In many forested regions around the world, the risk of severe drought is increasing ([Trenbert et al. 2014](https://www.nature.com/articles/nclimate2067)), even in conjunction with increasing precipitation ([IPCC 2014](https://www.cambridge.org/core/books/climate-change-2014-impacts-adaptation-and-vulnerability-part-b-regional-aspects/036A899BD52861D61B0D519C5F2B9334)). Global change-type drought has been affecting forests worldwide ([Allen et al. 2015](https://doi.org/10.1016/j.foreco.2009.09.001)), and it is expected that future climate change-driven droughts will severely impact forests around the world ([Allen et al. 2015](https://doi.org/10.1016/j.foreco.2009.09.001); **REFS**). Larger trees tend to suffer more (e.g., [@bennett_larger_2015]; Stovall et al. 2019), resulting in disproportionate impacts on forest C storage ([Meakem et al. 2018](https://nph.onlinelibrary.wiley.com/doi/10.1111/nph.14633)). As a result, forest drought responses stand to strongly impact forest feedbacks to climate change (**REFS**), yet accurate characterization of drought responses remains a modeling challenge (**REFS**)-- in part because some of the mechanisms underlying drought responses remain unclear. Understanding forest responses to drought requires increased functional understanding of how tree size, microhabitat, and species' traits jointly confer individual-level vulnerability or resistance, and the extent to which their influence is consistent across droughts.

@@

**A final fundamental question regarding forest responses to drought is whether tree size and species' traits have similar influence across droughts, or whether drought variability in factors such as severity, duration, and timing interact with tree size and traits such that different components of the community respond differently to different droughts.** No two droughts are the same, and tree growth responses vary with drought characteristics such as timing and atmospheric demand ([D'Orangeville et al. 2018](https://onlinelibrary-wiley-com.smithsonian.idm.oclc.org/doi/full/10.1111/gcb.14096)). However, we are not aware of any studies that compare how tree size and species' traits mediate growth responses across droughts. (**BUT ARE WE MISSING SOMETHING??**) While tree-ring studies provide long-term records of tree responses to multiple droughts (e.g., [@lloret_components_2011];  D'Orangeville et al. 2018 **REFS**), these don't test for differential trait effects across droughts (D'Orangeville et al. 2018) and generally focus on species-level responses, which preclude consideration of the roles of tree size and microenvironment. The ecological studies that have shaped our understanding of the role of tree size and microenvironment in forest drought responses generally examine only a single drought and tend to focus disproportionately on extreme droughts with dramatic impacts (*e.g.*, [Allen et al. 2015](https://doi.org/10.1016/j.foreco.2009.09.001); [@bennett_larger_2015]; Stovall; **MORE REFS**). Thus, our knowledge of forest responses to more modest but frequent droughts--e.g., those with historical return intervals on the order of a decade--remains more limited. While the tendency for larger trees to suffer more definitely predominates [@bennett_larger_2015], there are exceptions (e.g., **REFS**). There is also evidence that the degree to which larger trees suffer more increases with the severity of drought conditions ([@bennett_larger_2015]; *Stovall et al. 2019 (I think!)*). [*Are there any studies showing interactions of drought type with traits?*] Thus, while we expect many of the fundamental mechanisms shaping drought responses to be universal, we have little undertanding of how tree size and traits interact with drought characteristics to result in differential responses across droughts.
(**PELTIER**)

@@

Although the physiological mechanisms discussed above lead to generally consistent directions of growth responses to tree height and hydaulic traits across droughts, indicating the universality of the underlying mechanisms, the relative importantance of the drivers varied widely across droughts, indicating an interaction with drought characteristics (Tables 4-5). [**BRIEFLY DISCUSS DIFFERENCES ACROSS DROUGHTS-- e.g., 1966 was preceded by two dry years, and height mattered most in this drought. ...**] Of course, site characteristics also define the nature of droughts, and comparisons of size and trait effects across sites would be of great value to elucidating the mechanisms through which these shape drought resistance (**check for refs**). Unfortunately, our ability to compare forest responses across droughts is hampered by the fact that droughts are rarely explicitly defined in ecological studies [@slette_how_2019]. Further research will be required to understand how drought characteristics interact with driver variables to shape tree growth responses.

>>>>

\newpage
### Introduction
Forests globally play a critical role in climate regulation ([Bonan 2008](https://science.sciencemag.org/content/320/5882/1444)), yet there remains enormous uncertainty as to how the terrestrial carbon (C) sink, which is dominated by forests, will respond to climate change ([Friedlingstein et al. 2006](https://journals.ametsoc.org/doi/full/10.1175/JCLI3800.1)). An important aspect of this uncertainty lies in responses to drought (**REF**). In many forested regions around the world, the risk of severe drought is increasing ([Trenbert et al. 2014](https://www.nature.com/articles/nclimate2067)), even in conjunction with increasing precipitation ([IPCC 2014](https://www.cambridge.org/core/books/climate-change-2014-impacts-adaptation-and-vulnerability-part-b-regional-aspects/036A899BD52861D61B0D519C5F2B9334)). Global change-type drought has been affecting forests worldwide ([Allen et al. 2015](https://doi.org/10.1016/j.foreco.2009.09.001)), and it is expected that future climate change-driven droughts will severely impact forests around the world ([Allen et al. 2015](https://doi.org/10.1016/j.foreco.2009.09.001); **REFS**). Larger trees tend to suffer more (e.g., [@bennett_larger_2015]; [Stovall et al. 2019]( https://doi.org/10.1038/s41467-019-12380-6)), resulting in disproportionate impacts on forest C storage ([Meakem et al. 2018](https://nph.onlinelibrary.wiley.com/doi/10.1111/nph.14633)). As a result, forest drought responses stand to strongly impact forest feedbacks to climate change (**REFS**), yet accurate characterization of drought responses remains a modeling challenge (**REFS**)-- in part because some of the mechanisms underlying drought responses remain unclear. Understanding forest responses to drought requires increased functional understanding of how tree size, microhabitat, and species' traits jointly confer individual-level vulnerability or resistance, and the extent to which their influence is consistent across droughts.

@@

**A final fundamental question regarding forest responses to drought is whether tree size and species' traits have similar influence across droughts, or whether drought variability in factors such as severity, duration, and timing interact with tree size and traits such that different components of the community respond differently to different droughts.** No two droughts are the same, and tree growth responses vary with drought characteristics such as timing and atmospheric demand ([D'Orangeville et al. 2018](https://onlinelibrary-wiley-com.smithsonian.idm.oclc.org/doi/full/10.1111/gcb.14096)). However, we are not aware of any studies that compare how tree size and species' traits mediate growth responses across droughts. (**BUT ARE WE MISSING SOMETHING??**) While tree-ring studies provide long-term records of tree responses to multiple droughts (e.g., [@lloret_components_2011];  D'Orangeville et al. 2018 **REFS**), these don't test for differential trait effects across droughts (D'Orangeville et al. 2018) and generally focus on species-level responses, which preclude consideration of the roles of tree size and microenvironment. The ecological studies that have shaped our understanding of the role of tree size and microenvironment in forest drought responses generally examine only a single drought and tend to focus disproportionately on extreme droughts with dramatic impacts (*e.g.*, [Allen et al. 2015](https://doi.org/10.1016/j.foreco.2009.09.001); [@bennett_larger_2015]; Stovall; **MORE REFS**). Thus, our knowledge of forest responses to more modest but frequent droughts--e.g., those with historical return intervals on the order of a decade--remains more limited. While the tendency for larger trees to suffer more definitely predominates [@bennett_larger_2015], there are exceptions (e.g., **REFS**). There is also evidence that the degree to which larger trees suffer more increases with the severity of drought conditions ([@bennett_larger_2015]; Stovall et al. 2019). [*Are there any studies showing interactions of drought type with traits?*] Thus, while we expect many of the fundamental mechanisms shaping drought responses to be universal, we have little undertanding of how tree size and traits interact with drought characteristics to result in differential responses across droughts.
(**PELTIER**)

@@

Although the physiological mechanisms discussed above lead to generally consistent directions of growth responses to tree height and hydaulic traits across droughts, indicating the universality of the underlying mechanisms, the relative importantance of the drivers varied widely across droughts, indicating an interaction with drought characteristics (Tables 4-5). [**BRIEFLY DISCUSS DIFFERENCES ACROSS DROUGHTS-- For example, 1966 was preceded by two dry years, and height had the strongest effect in this drought, consistent with the finding that height becomes a stronger predictor of mortality as drought duration and intensity increase (Stovall et al. 2019). ...**] Of course, site characteristics also define the nature of droughts, and comparisons of size and trait effects across sites would be of great value to elucidating the mechanisms through which these shape drought resistance (**check for refs**). Unfortunately, our ability to compare forest responses across droughts is hampered by the fact that droughts are rarely explicitly defined in ecological studies [@slette_how_2019]. Further research will be required to understand how drought characteristics interact with driver variables to shape tree growth responses.
